---
title: 'Homework 3'
author: "Rocky Mazorow (Section 102)"
date: "4/4/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer after each code chunk.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed worksheet on D2L under the appropriate Homework folder.

***
1. Medical research has shown that repeated interval for extensions beyond 20 degrees increases the risk of wrist and hand injuries. Each of 24 students at Cornell University used a proposed new computer mouse design, and while using the mouse, each student's wrist extension was recorded. 

```{r}
wrist <- data.frame(ID = 1:24,
          extension = c(27, 28, 24, 26, 27, 25, 25, 24, 24, 24, 25, 28, 22, 25, 24, 28, 27, 26, 31, 25, 28, 27, 27, 25))
```

  i) Compute the mean and standard deviation of the wrist extensions data above. Describe the sample using these summaries.  
 
  ii) Create a histogram of the wrist extensions. Is the normality assumption reasonable?  
  
  iii) Use the data to estimate the mean wrist extension for people using this new mouse design using a 97\% confidence interval. Provide an interpretation of this interval.  
  
  
  iv) Use the data to test the hypothesis that the mean wrist extension for people using this new design is greater than 20 degrees. Use alpha = 0.01 (1\%) level of significance.  
  
  
  
### Code chunk
```{r} 
# Part A: Compute the mean and standard deviation of the wrist extensions data above. Describe the sample using these summaries.  
xBar <- mean(wrist$extension)
sigma <- sd(wrist$extension)
xBar
sigma

# Part B: Create a histogram of the wrist extensions. Is the normality assumption reasonable?  
hist(wrist$extension)  

# Part C: Use the data to estimate the mean wrist extension for people using this new mouse design using a 97% confidence interval. Provide an interpretation of this interval.  
n <- 24
margin <- qnorm(0.985)*sigma/sqrt(n)
lBound <- xBar - margin
uBound <- xBar + margin
lBound
uBound

# Part D: Use the data to test the hypothesis that the mean wrist extension for people using this new design is greater than 20 degrees. Use alpha = 0.01 (1%) level of significance. 
# Null  hypothesis is that µ ≤ 20
mu <- 20
t = (xBar - mu)/(sigma/sqrt(n))
pval <- pt(t, df=n-1, lower.tail=FALSE)
pval

```
### Code Analysis
i) The mean of extension angle is 25.92 degrees with a standard deviation of 1.954 degrees.
ii) The graph is not bell-shaped or symmetric about the mean. Therefore, the normality assumption is not reasonable.
iii) We are 97% confident that the mean wireless percentage is between 25.05 and 26.78 degrees.
iv) Since this p-value (1.447*10^-13) is less than our significance level (0.01), we reject the null hypothesis. Therefore, the mean wrist extension for people using this new design is greater than 20 degrees.
&nbsp;  
&nbsp; 

2. Recall the Going Wireless data first mentioned Week 2 of this class. The article Going Wireless (AARP Bulletin, June 2009) reported the estimated percentage of house-holds with only wireless phone service (no land line) for the 50 U.S. states and the District of Columbia. In the accompanying data table, each state was also classified into one of three geographical regions—West (W), Middle states (M), and East (E). Consider only the variable `Wireless` in this data.

```{r}
wireless.data <- read.csv("https://goo.gl/72BKSf", header = TRUE)
str(wireless.data)
```

  i) Compute the mean and standard deviation of the wireless data above. Describe the sample using these summaries.  
 
  ii) Create a histogram of the wireless data. Is the normality assumption reasonable?  
  
  iii) Use the data to estimate the mean wireless percentage per state using a 90\% confidence interval. Provide an interpretation of this interval.  
  
  
  iv) Use the data to test the hypothesis that the mean  wireless percentage per state is less than 17. Use alpha = 0.05 (5\%) level of significance.  
  
  
  
### Code chunk
```{r} 
# Part A: Compute the mean and standard deviation of the wireless data above. Describe the sample using these summaries.  
xBar <- mean(wireless.data$Wireless)
sigma <- sd(wireless.data$Wireless)
xBar
sigma

# Part B: Create a histogram of the wireless data. Is the normality assumption reasonable?  
hist(wireless.data$Wireless)  

# Part C: Use the data to estimate the mean wireless percentage per state using a 90% confidence interval. Provide an interpretation of this interval.  
n <- 31
margin <- qnorm(0.95)*sigma/sqrt(n)
lBound <- xBar - margin
uBound <- xBar + margin
lBound
uBound

# Part D: Use the data to test the hypothesis that the mean wireless percentage per state is less than 17. Use alpha = 0.05 (5%) level of significance. 
# Null  hypothesis is that µ ≥ 17
mu <- 17
t = (xBar - mu)/(sigma/sqrt(n))
pval <- pt(t, df=n-1)
pval

```
### Code Analysis
i) The mean of wireless users is 14.82% with a standard deviation of 5.344%.
ii) Although this data is more normal than Problem 1 (more bell-shaped and symmetrically distributed), I would still say the normality assumption is not reasonable due to the tail between 25 and 30.
iii) We are 90% confident that the mean wireless percentage is between 13.23% and 16.39%.
iv) Since this p-value (0.015) is less than our significance level (0.05), we reject the null hypothesis. Therefore, the mean wireless percentage per state is less than 17.
&nbsp;  
&nbsp; 