---
title: 'Worksheet 5'
author: "Rocky Mazorow (Section 102)"
date: "4/12/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer after each code chunk.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed worksheet on D2L under the appropriate Homework folder.

***
1.  While campaigning for higher political office during a recent election, a certain candidate claimed that \at least 75% of voters want the country to end all foreign aid to all countries." When a polling organization conducted a survey to investigate this claim, they found that 242 out of a sample of n = 346 expressed agreement with the statement. At a level of $\alpha = 0.02$, use the five-step framework to test
$H_0 : p \ge 0.75$ against $H_1 : p < 0.75$. What is the p-value? 


### Code chunk
```{r} 
# Step 0: Gather sample data
x <- 242
n <- 346 
p0 <- 0.75
p <- x/n

# Step 1: H_0: p >= 0.75 and H_1: p < 0.75

# Step 2: Significance level
alpha <- 0.02

# Step 3: Calculate test statistic z
z <- (p-p0)/(sqrt((p0*(1-p0))/n))
z

# Step 4: Calculate p-value of z
pval <- pnorm(z)
pval

# Alternate way
ptest <- prop.test(x,n, conf.level = 0.98, p = p0, alternative ='l', correct = FALSE)
ptest

```

### Code Analysis
Step 5: Draw a conclusion  
Since p-value (0.015) < alpha (0.02), we reject $H_0$.  
There is sufficient evidence that at least 75% of voters want the country to end all foreign aid to all countries.
&nbsp;  
&nbsp; 

2. After the 2012 US Presidential election in which Barack Obama was reelected for his second term in office, there was considerable discussion about the "gender gap" that existed between male and female voters in terms of their choice of candidate.Gallup; a survey group explored this issue after the election by conducting a poll of voting behavior among n=2551 voters. The results of their finding based on an exit poll is presented in the table below. At a level of $\alpha = 0.05$, use the five-step framework to test the null hypothesis that the proportion of women and men voting for Obama in 2012 is the same. 


```{r,echo=FALSE}
d <- cbind(Name = c("Obama","Romney","Totals"), Men = c(558,658,1214), Women = c(749,588,1337), Totals=c(1307,1244,2551))
# at most 4 decimal places
knitr::kable(d, digits = 4)
```
  


### Code chunk
```{r} 
# Step 0: Gather sample data
n1 <- 1214   # men
x1 <- 558    
n2 <- 1337   # women
x2 <- 749    
p1 <- x1/n1
p2 <- x2/n2

# Step 1: H_0: p1=p2 and H_1: p1!=p2

# Step 2: Significance level
alpha <- 0.05

# Step 3: Calculate test statistic z
p <- (p1*n1 + p2*n2)/(n1+n2)
z <- (p1-p2)/sqrt(p*(1-p)*(1/n1 + 1/n2))
z

# Step 4: Calculate p-value of z
pval <- pnorm(abs(z),lower.tail=FALSE)*2
pval

# Alternate way
x <- c(x1,x2)
n <- c(n1,n2)
ttest <- prop.test(x,n,conf.level = 0.95,correct = FALSE)
ttest
```

### Code Analysis
Step 5: Draw a conclusion  
Since p-value ($3.87*10^{-7}$) < alpha (0.05), we reject $H_0$.  
There is sufficient evidence that the proportions of women and men voting for Obama in 2012 is not the same.   
