---
title: 'Worksheet 1'
author: "Rocky Mazorow (Section 102)"
date: "2/2/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed worksheet on D2L under the worksheet folder.

***

1. Create a vector of three elements `(2,5,3)` and name that vector `vec.a`. Create a second vector, `vec.b`, that contains `(1,3,5)`. 
    a. Add these two vectors together and name the result `vec.c`.
    b. Create a vector, named `vec.d`, that contains only two elements `(5,7)`. 
    Add this vector to `vec.a`. What is the result? What is the warning message that `R` gives you? (look up the recycling rule using `Google`)? 
    c. Next add 5 to the vector `vec.a`. What is the result and what did `R` do? Why doesn’t it give you a warning message similar to what you saw in the previous problem?

## Code chunk
```{r,} 
# Insert your code for this question after this line

# Part A: Create vectors A and B with 3 elements. Store their sum in vector C  
vec.a <- c(2,5,3)
vec.b <- c(1,3,5)
vec.c <- vec.a + vec.b

# Part B: Create vector D with 2 elements. Add D to A. Report warning.
vec.d <- c(5,7)
vec.a + vec.d
    # Generates warning "longer object length is not a multiple of shorter object length"
    # R requires the smaller vector to have the same number of elements or multiples of the larger to evenly add across

# Part C: Add 5 to A. Report result and explain why there isn't a warning.
vec.a + 5
    # Result: [1]  7 10  8
    # Since 5 is only one element (multiple of 3), it can be added to each element

```


2. Generate the vector of even numbers `{2, 4, 6, . . . , 20}`
    a. Using the `seq()` function and
    b. Using the `a:b` shortcut and some subsequent algebra. *Hint: Generate the vector 1-10 and then multiple it by 2.*

```{r}
# Insert your code for this question after this line

# Part A: Generate vector of even numbers using seq()
vec.e1 <- seq(from = 2, to = 20, by = 2)

# Part B: Generate vector of even numbers using : and algebra
vec.e2 <- 1:10 * 2

```
  


3. The vector `letters` is a built-in vector to `R` and contains the lower case English alphabet.
    a. Extract the 9th element of the `letters` vector.
    b. Extract the sub-vector that contains the 9th, 11th, and 19th elements.
    c. Extract the sub-vector that contains everything except the last two elements.

```{r}
# Insert your code for this question after this line

# Part A: Generate vector of even numbers using seq()
vec.e1 <- seq(from = 2, to = 20, by = 2)

# Part B: Generate vector of even numbers using : and algebra
vec.e2 <- 1:10 * 2

# Problem 3
# Part A: 9th element of built in 'letters' vector
letters[9]

# Part B: sub-vector of 'letters' vector containing the 9th, 11th, and 19th elements
letters[ c(9,11,19) ]

# Part C: sub-vector of 'letters' vector containing everything but last two elements
letters[ c(-25, -26)]            # or letters[1:(length(letters)-2)]

```