---
title: 'Homework 1'
author: "Rocky Mazorow (Section 102)"
date: "2/9/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed worksheet on D2L under the worksheet folder.

***

1. Create a matrix whose entries are all prime numbers below 30.  
      i)  Create this matrix using the function `matrix()` with five rows. Save the matrix as `P`.  
      ii) Extract the second and third row out of `P`.  
      iii) Extract the entry in the fourth row and first column of `P`.  
      iv) Generate the transpose of the matrix using the function `t()`. What is the new dimension of this matrix.  

### Code chunk
```{r}
# Insert your code for this question after this line

# Part A: Create a matrix P whose entries are all prime numbers below 30.   
cells <- c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29)
P <- matrix(cells, nrow=5, ncol=2)
dim(P)

# Part B: Extract the second and third row out of `P`.
P[ c(2,3), ]

# Part C: Extract the entry in the fourth row and first column of `P`.
P[4,1]

# Part D: Generate the transpose of the matrix using the function `t()`. 
#         What is the new dimension of this matrix?
T <- t(P)
dim(T)

```

2. Create and manipulate a data frame.
      i) Create a `data.frame` named `my.trees` that has the following columns:  
          - `Girth = c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0)`  
          - `Height = c(70, 65, 63, 72, 81, 83, 66)`  
          - `Volume = c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)`  
      ii) Extract the fifth observational unit.  
      iii) Extract the `Girth` column referring to it by name.  
      iv) Print out a data frame of all the observations except for the last observation.  


### Code chunk
```{r}
# Insert your code for this question after this line

# Part A: Create a data.frame named `my.trees` that has the following columns:  
#         Girth = c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0) 
#         Height = c(70, 65, 63, 72, 81, 83, 66) 
#         Volume = c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)
Girth <- c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0) 
Height <- c(70, 65, 63, 72, 81, 83, 66) 
Volume <- c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)
my.trees <- data.frame(Girth, Height, Volume)

# Part B: Extract the fifth observational unit.  
my.trees[5,]

# Part C: Extract the `Girth` column referring to it by name. 
my.trees["Girth"]

# Part D: Print out a data frame of all the observations except for the last observation.
my.trees[c(1:(nrow(my.trees)-1)),]

```

3. The popular `iris` data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica. The `iris` data set is included with every R installation.  
      i) Check the structure of the`iris` data.   
      ii) How many variables and observations are in the data set?   
      iii) Which variables are `numeric` type?  
      iv) Display the first 4 rows of the data.  



### Code chunk
```{r} 
# Insert your code for this question after this line

# Part A: Check the structure of the`iris` data.   
str(iris)

# Part B: How many variables and observations are in the data set?   
#     There are 5 variables and 150 observations

# Part C: Which variables are `numeric` type?  
#     The numeric variables are Sepal.Length, Sepal.Width, Petal.Length, Petal.Width

# Part D: Display the first 4 rows of the data. 
iris[1:4,]

```



  
4.  You will work with the passenger Titanic data found on this [link](<https://goo.gl/NHb1Pg>).   
      i) Read this data directly from the web using its URL.  
      ii) Read this data by saving it first into your working directory. Check your working directory using `getwd()` or change its location using the steps covered in the lesson.  
      iii) How many passengers and what variables are in the data set?   
      iv) Display the first 5 rows of the data.  

### Code chunk
```{r}
# Insert your code for this question after this line

# Part A: Read Titanic data directly from the web: <https://goo.gl/NHb1Pg> 
titanic.data1 <- read.csv("https://goo.gl/NHb1Pg", header = TRUE)

# Part B: Read this data by saving it first into your working directory. 
#         Check your working directory using `getwd()` or change its location using the steps covered in the lesson.  
getwd()
titanic.data2 <- read.csv("/Users/rmazorow/Library/CloudStorage/OneDrive-MarquetteUniversity/School Work/MSSC 5931/Intro-to-R-Prog/Titanic_Data.csv", header = TRUE)

# Part C: How many passengers and what variables are in the data set?   
str(titanic.data1)
#   There are 1316 passengers
#   The variables are X, class, age, sex, and survived

# Part D: Display the first 5 rows of the data.  
titanic.data1[1:5, ]
```
