---
title: 'Worksheet 3'
author: "Rocky Mazorow (Section 102)"
date: "3/11/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed worksheet on D2L under the worksheet folder.

***

1. Consider the population of CEO salaries in the lesson this week.   

    i) Select a random sample of 40 CEO's and compute the sample mean salary and sample standard deviation salary.  
    
    ii) Use the `sample()` function and the `for loop` in the lecture note to compute the sample mean of 100 randomly selected samples.  Plot the histogram of these 100 sample means.  
    
    iii) Compute the average and standard deviation of these 100 sample means. Are the values what you expect from the properties of the sampling distribution?

### Code chunk
```{r} 
# Setting seed
set.seed(123456)

# Load data set containing the annual salary (in $Million) information of 497 CEO’s in 2012
ceo.data <- read.csv("/Users/rmazorow/Library/CloudStorage/OneDrive-MarquetteUniversity/School Work/MSSC 5931/Intro-to-R-Prog/data/CEO_Salary_2012.csv", header = TRUE)
mu <- mean(ceo.data$Annual.pay)
sigma <- sd(ceo.data$Annual.pay)

# Part A: Select a random sample of 40 CEO's and compute the sample mean salary and sample standard deviation salary.
s1  <- sample(ceo.data$Annual.pay,40)
mean(s1)
sd(s1)

# Part B: Use the `sample()` function and the `for loop` in the lecture note to compute the sample mean of 100 randomly selected samples. Plot the histogram of these 100 sample means.
nn <- 100
ss <- matrix(NA, nrow=40, ncol=nn)
for (j in 1:nn) ss[,j] <- sample(ceo.data$Annual, 40)
sn <- apply(ss, 2, mean)

hist(sn, main="Sample Mean (Size n= 40) Salary Distribution", xlab="Sample Mean Salary ($Mil)")

# Part C: Compute the average and standard deviation of these 100 sample means. Are the values what you expect from the properties of the sampling distribution?
mn2 <- mean(sn)
sd2 <- sd(sn)

#    Compare true values to sample values
c(mu, mn2)    
c(sigma/sqrt(40), sd2)
#    I expected the sample mean to be comparable to the true mean. Since 40 samples is close to a sufficiently large sample size (n=30), I expected the sample standard deviation to be close to, if not a little less, than the true standard deviation. The results support both of these assumptions.


```

2. Rockwell hardness of pins of a certain type is known to have a mean value of 50 and standard deviation of 1.2.

    i) If the distribution is normal, what is the probability that the sample mean hardness for a random sample of 9 pins is at least 51?     

    ii) Without assuming population normality, what is the  (approximate) probability that the sample mean hardness for a random sample of 40 pins is at least 51?
  

### Code chunk
```{r} 
# Part A: If the distribution is normal, what is the probability that the sample mean hardness for a random sample of 9 pins is at least 51?     
pnorm(q = 51, mean = 50, sd = 1.2/sqrt(9), lower.tail=FALSE)

# Part B: Without assuming population normality, what is the (approximate) probability that the sample mean hardness for a random sample of 40 pins is at least 51?
pnorm(q = 51, mean = 50, sd = 1.2/sqrt(40), lower.tail=FALSE)
```


3. Suppose that a random sample of size 64 is to be selected from a population with mean 40 and standard deviation 5.

    i) What are the mean and standard deviation of the sampling distribution of the sample mean?
    
    ii) What is the approximate probability that the sample mean will be within 0.5 of the population mean?  
    
### Code chunk
```{r} 
# Part A: What are the mean and standard deviation of the sampling distribution of the sample mean?
#    Sample is expected to have mean of 40 and standard deviation of 5. 
r  <- rnorm(1000, mean = 40, sd = 5)
s3 <- sample(r, 64)
mean(s3)
sd(s3)

# Part B: What is the approximate probability that the sample mean will be within 0.5 of the population mean? 
pnorm(40.5, mean=40, sd=5/sqrt(64))-pnorm(39.5, mean=40, sd=5/sqrt(64))
```