---
title: 'Worksheet 4'
author: "Rocky Mazorow (Section 102)"
date: "3/25/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed worksheet on D2L under the worksheet folder.

***
### Investgating the effect of change in sample 
1. The sample mean weight of a random sample of 25 turtles is 300 with a population standard deviation of 18.5. Investigate the effect of sample size on confidence interval by changing the sample size from

    i) 25 to 100 and finding a 95% confidence interval for the population mean.
    
    ii) 25 to 10 and finding a 95% confidence interval for the population mean.
    
    

### Code chunk
```{r} 
# Original sample
n <- 25
xBar <- 300
sigma <- 18.5
margin <- qnorm(0.975)*sigma/sqrt(n)
lBound <- xBar - margin
uBound <- xBar + margin
lBound
uBound 

# Part A: 25 to 100 and finding a 95% confidence interval for the population mean.
n <- 100
margin <- qnorm(0.975)*sigma/sqrt(n)
lBound <- xBar - margin
uBound <- xBar + margin
lBound
uBound 

# Part B: 25 to 10 and finding a 95% confidence interval for the population mean.
n <- 10
margin <- qnorm(0.975)*sigma/sqrt(n)
lBound <- xBar - margin
uBound <- xBar + margin
lBound
uBound 

# As the sample size increases, the confidence interval narrows.
# As the sample size decreases, the confidence interval widens.

```

2. For a class project, a political science student at a large university wants to estimate the percent of students that are registered voters. He surveys 500 students and finds that 300 are registered voters. Compute a 90% confidence interval for the true percent of students that are registered.
  

### Code chunk
```{r} 
n <- 500
p <- 300/n
margin <- qnorm(0.95)*sqrt(p*(1-p)/n)
lBound <- p - margin
uBound <- p + margin
lBound * 100
uBound * 100

# We are 90% confidence that the population of registered voters is between 56.4% and 63.6%.

```


